---
title: "Deliverable 3"
author: "Jay Gohner"
date: "December 13, 2019"
output: html_document
---


```{r GlobalOptions, echo=FALSE}
options(knitr.duplicate.label = 'allow')
```

In this part we bring in all of the data and graphs from the first two phases so that we may discuss the operationalization of the data
and the summary of what we have found and what could be done with this information.

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}
include <- function(library_name){
  if( !(library_name %in% installed.packages()) )
    install.packages(library_name) 
  library(library_name, character.only=TRUE)
}
include("tidyverse")
include("knitr")
purl("deliverable2.rmd", output = "part2.r")
source("part2.r")
options(scipen=999)
```

## PEER REVIEW

My previous work was peer reviewed by Kyle Kirk and this is what he had to say.

<h1>Modeling</h1>

There was a lot of work that went into building the tables for this model. The model had a very good p-value so the model is on to
something. The graph does not represent the model well so that would need to be fixed.

<h1>R Proficiency</h1> 

Upon reviewing the code it was clear that there was a strong understanding of R and some creative custom functions. I worked with similar
education data and know the process well.

<h1>Communication</h1>
All of the R code chunks had very thourough documentation and presented a clear picture of what was being done and why. I feel with no
knowledge of R I could still understand the process that was undergone for this project.

<h1>Critical Thinking</h1>
The project looked to show a correalation between rising levels of adults obtaining higher levels of education and a decrease in incidents
of violent crime. It is important topic to be studied and while a relatively simple correalation it is still an important one.

<h1>Summary</h1>
Overall the project seems to accomplish it's goals of finding a correalation between the two datasets. There are obviously many other
factors that will contribute to this so that would be worth exploring in the future.


<h3>Cross-Validation</h3>
While the data I have does show a positive correaltion like I had hoped I believe more data is needed to back up my study. The problem is
finding data that matches state year and demographics you are looking for. So much is behind a paywall or just plain unavailable.

<h3>Operationalization</h3>
This data shows that when there is an increase in the amount of adults obtaining a higher level of education that there will be a decrease
in the number of violent crimes in a state. As Kyle mentioned in my peer-review there are definitely other factors that will affect crime
levels going down as well. I personally think the next step would be to somehow gather data on the quality of education in a state of study
and the overall quality of social services and resources available to residents in that state. With this data perhaps more funding to
education and extra-curricular actvities could applied to areas with higher rates of violence per capita.

The things I worry about with a study like this is in our current hyper-partisan political state is that certain groups could be excluded
and not benefit from the increase in funding.




