---
title: "Deliverable 3"
author: "Jay Gohner"
date: "December 13, 2019"
output: html_document
---


```{r GlobalOptions, echo=FALSE}
options(knitr.duplicate.label = 'allow')
```

In this part we bring in all of the data and graphs from the first two phases so that we may discuss the operationalization of the data
and the summary of what we have found and what could be done with this information.

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, results='hide'}
include <- function(library_name){
  if( !(library_name %in% installed.packages()) )
    install.packages(library_name) 
  library(library_name, character.only=TRUE)
}
include("tidyverse")
include("knitr")
purl("deliverable2.rmd", output = "part2.r")
source("part2.r")
options(scipen=999)
```

## PEER REVIEW

My previous work was peer reviewed by Kyle Kirk and this is what he had to say.

<h1>Modeling</h1>

There was a lot of work that went into building the tables for this model. The model had a very good p-value so the model is on to
something. The graph does not represent the model well so that would need to be fixed.

<h1>R Proficiency</h1> 

Upon reviewing the code it was clear that there was a strong understanding of R and some creative custom functions. I worked with similar
education data and know the process well.

<h1>Communication</h1>
All of the R code chunks had very thourough documentation and presented a clear picture of what was being done and why. I feel with no
knowledge of R I could still understand the process that was undergone for this project.

<h1>Critical Thinking</h1>
The project looked to show a correalation between rising levels of adults obtaining higher levels of education and a decrease in incidents
of violent crime. It is important topic to be studied and while a relatively simple correalation it is still an important one.

<h1>Summary</h1>
Overall the project seems to accomplish it's goals of finding a correalation between the two datasets. There are obviously many other
factors that will contribute to this so that would be worth exploring in the future.

<h2>REWORK</h2>
I had to make quite a few name chages in phase 1 to make an actual graph of the education data. In phase 2 I had to make new columns of per
capita data for both diploma levels and violent crime totals to make a model that could give me interpretable results.

<h3>Cross-Validation</h3>
While the data I have does show a positive correaltion like I had hoped I believe more data is needed to back up my study. The problem is
finding data that matches state year and demographics you are looking for. So much is behind a paywall or just plain unavailable.

<h3>Operationalization</h3>
Unfortunately this data and model do not give a very accurate indication of the rates of violent crime decreasing when there is an increase
in the rates of adults obtaining higher levels of education. I believe the problem primarily relys on the fact that there are likely many
factors that can contribute to the decrease in violents rates of crimes in an adult population. As Kyle mentioned in my peer-review there 
are definitely other factors that will affect crime levels going down as well. To further improve this study there are several areas where 
I believe data should be collected to improve the accuracy of the model.

Collecting data on the social programs available in a state, the amount of funding directed to the those services and the overall
satisfaction the population of the target area has regarding said programs. I believe the social safety net programs we have for those
with less advantage help to curb desperation which I believe fuels rates of violence.

I would also like to see the GDP and unemployment rate of a state added to the model. This factors would also play a role in the quality
of peoples lives which I believe would have a positive effect on decreasing rates of violent crime.

And lastly I think it would be important to gather data on the quality of education in a state of study. With this data perhaps more funding to education and extra-curricular actvities could be applied to areas with higher rates of violence per capita.

The things I worry about with a study like this is in our current hyper-partisan political state is that certain groups could be excluded
and not benefit from the increase in funding.




